---
title: "gapminder tidyverse examples"
---

Data from [Gapminder](https://www.gapminder.org/): the GM foundation

> promotes sustainable global development and achievement of the United Nations Millennium Development Goals by increased use and understanding of statistics and other information about social, economic and environmental development at local, national and global levels ...

Loading packages and reading data:

```{r setup}
library(tidyverse)
library(plotly)
source("http://ms.mcmaster.ca/bolker/misc/gapminder_funs.R")
```

Creates a `gm_ind` object in the workspace.

```{r}
## interactive sheet (or via RStudio "spreadsheet" icon)
View(gm_ind)
(gm_ind
  %>% filter(str_detect(indicator,"HIV"))
)
(gm_ind
  %>% filter(subcategory=="HIV")
)
(gm_ind
  %>% filter(subcategory %in% c("HIV","TB"))
)
(gm_ind
  %>% filter(str_detect(subcategory,"(HIV|TB)"))
)
(gm_ind
  %>% filter(str_detect(subcategory,"(HIV|TB)"))
  %>% select(-URL)
)
```

Get data and clean it up:

```{r get_data}
hiv1 <- get_data("Adults with HIV")
hiv2 <- (hiv1
  %>% rename(country="Estimated HIV Prevalence% - (Ages 15-49)")
)
hiv3 <- (hiv2
  %>% tidyr::gather(year,prevalence,-country, convert=TRUE)
  %>% drop_na()
)
```

```{r}
## ggplot package already loaded
theme_set(theme_bw())
(ggplot(hiv3,aes(year,prevalence,colour=country))
  + geom_line()
  + geom_point()
  + theme(legend.position="none")
  + scale_y_log10()
)
```

Merge with population data:

```{r}
pop1 <- get_data("Population, total")
pop2 <- (pop1
  %>% rename(country="Total population")
  %>% tidyr::gather(year,population,-country, convert=TRUE)
  %>% drop_na()
)
hiv4 <- (hiv3
  %>% left_join(pop2)
  %>% left_join(region_tab)
)
```

```{r}
arrange(hiv4,desc(population)) %>% tail()
```

```{r}
plot(ecdf(log10(hiv4$population)))
```

```{r}
hiv5 <- filter(hiv4, population>1e7)
gg2 <- (ggplot(hiv5,aes(year,prevalence,colour=region,label=country))
  + geom_line(aes(group=country))
  + geom_point()
)
ggplotly(gg2)
```

```{r}
hiv6 <- (hiv3
  %>% filter(year==2011)
  %>% select(country, prevalence)
  %>% rename(region="country",value="prevalence")
  %>% mutate(region=tolower(region))
)
country_choropleth(df=hiv6)
```
